---
title: "Experiments CMI & HiCS"
author: "Ines Machinek"
date: "December 6th, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(data.table)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
read_chunk("generatePlots.R")
```

##  Comparing CMI, HiCS and GMD Subspace Search using real-world datasets

### Overview
This document contains the results of a comparison of the algorithms CMI, HiCS & GMD.
For further documentation, see:
<ul>
  <li>
    Nguyen, Hoang, Vreeken, Jilles, Mueller, Emmanuel, Keller, Fabian and Klemens Boehm. 2013.
    "CMI: An Information-Theoretic Contrast Measure for Enhancing Subspace Cluster and Outlier Detection"
  </li>
  <li>
    Keller, Fabian, Emmanuel Mueller, and Klemens Boehm. 2012. "HiCS: High Contrast Subspaces for Density-Based Outlier Ranking." In 2012 IEEE 28th   
    International Conference on Data Engineering, 1037-48. IEEE.
  </li>
</ul>
CMI and HiCS algorithms were implemented in inline c++ in R, using the Rcpp-package.
Experiments ran on Intel core i7-4600U CPU and 8 GB RAM.

For implementation details see [R-Subcon](https://github.com/holtri/Rsubcon).
<br>
<br>

### Datasets
The datasets for testing were:

```{r echo = FALSE, results = 'asis'}

Datasets <- c("Ann Thyroid", "Diabetes", "Glass", "Ion", "Lympho", "Pendigits", "Segment")
Rows <- c(3772, 768, 214, 351, 148, 6737, 2013)
Dimensions <- c(6, 8, 7, 32, 18, 16, 19)

df <-data.frame(Datasets,Rows,Dimensions)

kable(df, caption = "Datasets - Overview")
```


Datasets can be downloaded here: 
[IPD, KIT Karlsruhe](http://www.ipd.uni-karlsruhe.de/~muellere/HiCS/)
<br>
<br>

### Run Experiments
1. Download the datasets into the *datasets* folder
2. Source the *HICSvsCMI_Experiments.R* script
3. Start experiment by calling **runExperiments**-function, for example:
```{r eval=FALSE}
  finalResult <- runExperiments(inputPath = "datasets", maxMinPts = 100, numCores=1, topkSearch = 500, topkOutput = 100)
```
Automatically a file *experimentResult.RData* will be written to the *results* folder.
If you want to reuse the same structure of the KNITR-File for other experiments, please make sure that you've got 1 RMD-Resultfile per dataset.

Results can be loaded by
```{r eval=FALSE}
load("results/experimentResult.RData")
```
<br>
<br>


### Results
The following results, unless otherwise noted, were observed after having run the algorithms with the parameters: <br>
* maxMinPts = 100   <br>
* numCores = 1      <br>
* topkSearch = 50   <br>
* topkOutput = 20   <br>

<br>
<br>


```{r, echo=FALSE, warning=FALSE}
# load data 
<<loadData>>
```

#### AUC on real world data 


```{r eval=T, echo=FALSE, warning=FALSE}
# AUC 
<<auc>>
```

<br>
<br>

#### Runtime

The duration of subspace search algorithm and LOF calculation were the following:

```{r, echo=FALSE, warning=FALSE}
# RUNTIME
<<runtime>>
``` 

#### Subspaces

One aspect we considered was the size of the subspaces which were found by the algorithms.

```{r eval=T, echo=FALSE, warning=FALSE}
# Mean on top20-subspaces:
<<meanSizeTop20SS>> 
``` 
<br>
<br>
With CMI subspaces with higher dimensions were found. It is even more obvious when you look at the Top5 Subspaces.

```{r eval=T, echo=FALSE, warning=FALSE}
# Mean on top5-subspaces:
<<meanSizeTop5SS>> 
```


Some examples of found subspaces:
```{r, echo=FALSE, warning=FALSE}
# Examples of top5-subspaces:
<<sampleSS>>
```


The size of the subspaces is much higher for CMI than for HiCs and for GMD. 
A reason is maybe the higher score a subspace gets by adding further dimensions. 

<br>
<br>
The global mean of subspace size is 

```{r eval=T, echo=FALSE, warning=FALSE}
#   global mean
  print(globalMean <- mean(subspaces$length))
```

Those are the mean sizes of subspaces per algorithm:

```{r eval=T, echo=FALSE, warning=FALSE}
#   mean per algorithm
<<meanSizealg>>
``` 



