---
title: "Experiments CMI & HiCS"
author: "Ines Machinek"
date: "December 6th, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(data.table)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
read_chunk("generatePlots.R")
```

##  Comparing CMI, HiCS and GMD Subspace Search using real-world datasets

### Overview
This document contains the results of a comparison of the algorithms CMI, HiCS & GMD.
For further documentation, see:
<ul>
  <li>
    Nguyen, Hoang, Vreeken, Jilles, Mueller, Emmanuel, Keller, Fabian and Klemens Boehm. 2013.
    "CMI: An Information-Theoretic Contrast Measure for Enhancing Subspace Cluster and Outlier Detection"
  </li>
  <li>
    Keller, Fabian, Emmanuel Mueller, and Klemens Boehm. 2012. "HiCS: High Contrast Subspaces for Density-Based Outlier Ranking." In 2012 IEEE 28th   
    International Conference on Data Engineering, 1037-48. IEEE.
  </li>
</ul>
CMI and HiCS algorithms were implemented in inline c++ in R, using the Rcpp-package.
Experiments ran on Intel core i7-4600U CPU and 8 GB RAM.

For implementation details see [R-Subcon](https://github.com/holtri/Rsubcon).
<br>
<br>

### Datasets
The datasets for testing were:

```{r echo = FALSE, results = 'asis'}

Datasets <- c("Ann Thyroid", "Diabetes", "Glass", "Ion", "Lympho", "Pendigits", "Segment")
Rows <- c(3772, 768, 214, 351, 148, 6737, 2013)
Dimensions <- c(6, 8, 7, 32, 18, 16, 19)

df <-data.frame(Datasets,Rows,Dimensions)

kable(df, caption = "Datasets - Overview")
```


Datasets can be downloaded here: 
[IPD, KIT Karlsruhe](http://www.ipd.uni-karlsruhe.de/~muellere/HiCS/)
<br>
<br>

### Run Experiments
1. Download the datasets into the *datasets* folder
2. Source the *HICSvsCMI_Experiments.R* script
3. Start experiment by calling **runExperiments**-function, for example:
```{r eval=FALSE}
  finalResult <- runExperiments(inputPath = "datasets", maxMinPts = 100, numCores=1, topkSearch = 500, topkOutput = 100)
```
Automatically a file *experimentResult.RData* will be written to the *results* folder.
If you want to reuse the same structure of the KNITR-File for other experiments, please make sure that you've got 1 RMD-Resultfile per dataset.

Results can be loaded by
```{r eval=FALSE}
load("results/experimentResult.RData")
```
<br>
<br>


### Results
The following results, unless otherwise noted, were observed after having run the algorithms with the parameters: <br>
* maxMinPts = 100   <br>
* numCores = 1      <br>
* topkSearch = 50   <br>
* topkOutput = 20   <br>

<br>
<br>


```{r, echo=FALSE, warning=FALSE}
# load data 

colSSVec <- c(1,2,36:55) # top 20 subspaces as vectors
colTop5SS <- c(1,2,16:20)   # top 5 subspaces as characters

load("results/Param_100_050_020_SSVec/experimentResult_ann_161129.RData")
annSS <-combinedResult[,colSSVec, with=FALSE]
annSSU <-annSS[!duplicated(annSS$algorithm)]
ann5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
tmp <- combinedResult 

# load("results/Param_100_050_020_SSVec/experimentResult_ann_161129.RData")
# breastSS <-combinedResult[,colSSVec, with=FALSE]
# breastU <-breastSS[!duplicated(breastSS$algorithm)]
# breast5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
# tmp <- rbind(tmp, combinedResult)

load("results/Param_100_050_020_SSVec/experimentResult_diabetes_161206.RData")
diabSS <-combinedResult[,colSSVec, with=FALSE]
diabSSU <-diabSS[!duplicated(diabSS$algorithm)]
diab5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
tmp <- rbind(tmp, combinedResult)

load("results/Param_100_050_020_SSVec/experimentResult_glass_161129.RData")
glassSS <-combinedResult[,colSSVec, with=FALSE]
glassSSU <-glassSS[!duplicated(glassSS$algorithm)]
glass5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
tmp <- rbind(tmp, combinedResult)

load("results/Param_100_050_020_SSVec/experimentResult_ion_161129.RData")
ionSS <-combinedResult[,colSSVec, with=FALSE]
ionSSU <-ionSS[!duplicated(ionSS$algorithm)]
ion5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
tmp <- rbind(tmp, combinedResult)

load("results/Param_100_050_020_SSVec/experimentResult_lymph_161206.RData")
lympSS <-combinedResult[,colSSVec, with=FALSE]
lympSSU <-lympSS[!duplicated(lympSS$algorithm)]
lymp5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
tmp <- rbind(tmp, combinedResult)

load("results/Param_100_050_020_SSVec/experimentResult_pen_161206.RData")
penSS <-combinedResult[,colSSVec, with=FALSE]
penSSU <-penSS[!duplicated(penSS$algorithm)]
pen5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
tmp <- rbind(tmp, combinedResult)

load("results/Param_100_050_020_SSVec/experimentResult_segmentn_161206.RData")
segmSS <-combinedResult[,colSSVec, with=FALSE]
segmSSU <-segmSS[!duplicated(segmSS$algorithm)]
segm5SS <-combinedResult[!duplicated(combinedResult[,colTop5SS, with=FALSE]),colTop5SS, with=FALSE]
combinedResult <- rbind(tmp, combinedResult)

```

#### AUC on real world data 


```{r eval=T, echo=FALSE, warning=FALSE}
# <<plotTest>>

# AUC 

combinedResult %>%   
select(dataset, algorithm, AUC_sum) %>%
  
  group_by(dataset, algorithm) %>%
  slice(which.max(AUC_sum)) %>%
 group_by(dataset, algorithm) %>%
  summarize(mean.AUC = mean(AUC_sum), sd.AUC = sd(AUC_sum)) %>%
  select(dataset, algorithm, mean.AUC, sd.AUC) %>%
  
  ggplot(aes(x = factor(dataset), y = mean.AUC, fill = algorithm)) +
  geom_bar(position = position_dodge(), stat="identity", width = 0.7) +
  geom_errorbar(aes(ymin=mean.AUC - sd.AUC, ymax=mean.AUC + sd.AUC), width=.2, position = position_dodge(.7)) +
  xlab(label = "Dataset") +
  ylab(label = "AUC") +
  scale_fill_grey(start = 0.1, end = 0.7)

```

<br>
<br>

#### Runtime

The duration of subspace search algorithm and LOF calculation were the following:

```{r, echo=FALSE, warning=FALSE}

# RUNTIME

  combinedResult %>%   
  select(dataset, algorithm, durationLOF, duationSS) %>%
  
  group_by(dataset, algorithm) %>%
  slice(which.max(sum(duationSS, durationLOF))) %>%
  group_by(dataset, algorithm) %>%
  
  mutate(runtime = sum(duationSS, durationLOF)) %>%
  
  ggplot(aes(x = factor(dataset), y = runtime, fill = algorithm)) +
  geom_bar(position = position_dodge(), stat="identity", width = 0.7) +
  xlab(label = "Dataset") +
  ylab(label = "Runtime in sec.") +
  scale_fill_brewer(palette = "Set2") 
  
``` 

#### Subspaces

One aspect we considered was the size of the subspaces which were found by the algorithms.

```{r eval=T, echo=FALSE, warning=FALSE}
# Mean on top20-subspaces:


ssv <- rbind(annSSU, diabSSU, glassSSU, ionSSU, lympSSU, penSSU, segmSSU)

 # gather first 
long_DF <- ssv %>% gather(name, subsp, subsp1V:subsp20V)


long_DF <- long_DF[order(long_DF$algorithm),]

# subspace length
lengthSS <- c(length(long_DF$subsp[[1]]))
#   .. add length as new column "length"
for (i in 2:length(long_DF$subsp)){
    lengthSS <- c(lengthSS, length(long_DF$subsp[[i]]))
}
subspaces <- mutate(long_DF, length = lengthSS) 

# mean per algorithm per dataset
algdsDistinct <- subspaces %>% select(algorithm, dataset) %>% distinct

dfMean <- data.frame()
for (i in 1:nrow(algdsDistinct)){
 
  tmpdf <- filter(subspaces, subspaces$algorithm == algdsDistinct$algorithm[i] & subspaces$dataset == algdsDistinct$dataset[i])             %>% select(length)
  
  tmpMean <- mean(tmpdf$length)
  # print(tmpMean)
  v <- data.frame(dataset = algdsDistinct$dataset[i], algorithm = algdsDistinct$algorithm[i], mean = tmpMean)
  dfMean  <- rbind(dfMean, v)

}


ggplot(dfMean, aes(factor(dataset), mean, fill=algorithm, x=dataset, y=mean, gg)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Set2") +
  labs(title ="Average size of Top20-Subspaces", y = "Mean size") 
``` 
<br>
<br>
With CMI subspaces with higher dimensions were found. It is even more obvious when you look at the Top5 Subspaces.

```{r eval=T, echo=FALSE, warning=FALSE}
# Mean on top5-subspaces:
top5Length <- filter(subspaces, subspaces$name == "subsp1V" | subspaces$name == "subsp2V" | subspaces$name == "subsp3V"
                | subspaces$name == "subsp4V" | subspaces$name == "subsp5V") 

dfMean5 <- data.frame()
for (i in 1:nrow(algdsDistinct)){
 
  tmpdf5 <- filter(top5Length, top5Length$algorithm == algdsDistinct$algorithm[i] & top5Length$dataset == algdsDistinct$dataset[i])             %>% select(length)
  
  tmpMean5 <- mean(tmpdf5$length)
  # print(tmpMean)
  v <- data.frame(dataset = algdsDistinct$dataset[i], algorithm = algdsDistinct$algorithm[i], mean = tmpMean5)
  dfMean5  <- rbind(dfMean5, v)
}

ggplot(dfMean5, aes(factor(dataset), mean, fill=algorithm, x=dataset, y=mean, gg)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Set2") +
  ggtitle("Average size of Top5-Subspaces")


```


Some examples of found subspaces:
```{r, echo=FALSE, warning=FALSE}
# Examples of top5-subspaces:

colNam <- c("Algorithm","Dataset","Subspace 1","Subspace 2","Subspace 3","Subspace 4","Subspace 5")

kable(glass5SS, caption = "Top-5-Subspaces - Glass-Dataset", col.names = colNam)
kable(ion5SS, caption = "Top-5-Subspaces - Ion-Dataset", col.names = colNam)
kable(pen5SS, caption = "Top-5-Subspaces - Pendigits-Dataset", col.names = colNam)

```


The size of the subspaces is much higher for CMI than for HiCs and for GMD. 
A reason is maybe the higher score a subspace gets by adding further dimensions. 

<br>
<br>
The global mean of subspace size is 

```{r eval=T, echo=FALSE, warning=FALSE}
#   global mean
  print(globalMean <- mean(subspaces$length))
```

Those are the mean sizes of subspaces per algorithm:

```{r eval=T, echo=FALSE, warning=FALSE}
#   mean per algorithm
meanAlg <- data.frame()
SSCMI <- filter(subspaces, subspaces$algorithm == "CMI")
meanCMI <- mean(SSCMI$length)
nL <- data.frame(algorithm = "CMI", mean = meanCMI)
meanAlg <- rbind(meanAlg, nL)

SSHics <- filter(subspaces, subspaces$algorithm == "HiCS")
meanHics <- mean(SSHics$length)
nL <- data.frame(algorithm = "HiCS", mean = meanHics)
meanAlg <- rbind(meanAlg, nL)

SSGMD <- filter(subspaces, subspaces$algorithm == "GMD")
meanGMD <- mean(SSGMD$length)
nL <- data.frame(algorithm = "GMD", mean = meanGMD)
meanAlg <- rbind(meanAlg, nL)

ggplot(meanAlg, mapping = aes(factor(algorithm), mean, x=algorithm, y=mean, gg)) +
geom_bar(stat="identity", position = "dodge") + 
  scale_color_brewer(palette = "Set2") +
  labs(title ="Average size of Top20-Subspaces per algorithm", y = "Mean size") 
``` 



