---
title: "Experiments CMI & HiCS"
author: "Ines Machinek"
date: "25 Oktober 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Comparing CMI and HiCS Subspace Search using real-world datasets

### Overview
This document contains the results of a comparison of the two algorithms CMI & HiCS.
For further documentation, see:
<ul>
  <li>
    Nguyen, Hoang, Vreeken, Jilles, M?ller, Emmanuel, Keller, Fabian and Klemens Bohm. 2013.
    "CMI: An Information-Theoretic Contrast Measure for Enhancing Subspace Cluster and Outlier Detection"
  </li>
  <li>
    Keller, Fabian, Emmanuel M?ller, and Klemens Bohm. 2012. "HiCS: High Contrast Subspaces for Density-Based Outlier Ranking." In 2012 IEEE 28th   
    International Conference on Data Engineering, 1037-48. IEEE.
  </li>
</ul>
Both algorithms were implemented in inline c++ in R, using the Rcpp-package.
Experiments ran on Intel core i7-4600U CPU and 8 GB RAM.


### Datasets
The datasets for testing were:

```{r echo = FALSE, results = 'asis'}

Datasets <- c("Thyroid", "Glass", "Ion", "Pendigits", "Segment", "Lympho")
Rows <- c(3772, 214, 351, 6737, 2013, 148)
Dimensions <- c(6, 7, 32, 16, 19, 18)

df <-data.frame(Datasets,Rows,Dimensions)
library(knitr)

kable(df, caption = "Datasets - Overview")
```


Datasets can be downloaded here: 
[http://www.ipd.uni-karlsruhe.de/~muellere/HiCS/]

### Run Experiments
1. Download the datasets into the *datasets* folder
2. Source the *HICSvsCMI_Experiments.R* script
3. Start experiment by calling **runExperiments**-function
```{r eval=FALSE}
  finalResult <- runExperiments(inputPath = "datasets", maxMinPts = 100, numCores=1, topkSearch = 500, topkOutput = 100)
```
Automatically a file *experimentResult.RData* will be written to the *results* folder.
Results can be loaded by
```{r eval=FALSE}
load("results/experimentResult.RData")
```

### Results
The following results were observed after having run the algorithms with the parameters: <br>
* maxMinPts = 100 <br>
* numCores=1 <br>
* topkSearch = 500 <br>
* topkOutput = 100 <br>


#### AUC
```{r, echo=FALSE}
Dataset = c("Thyroid","Thyroid","Glass","Glass","Ion", "Ion", "Pendigits","Pendigits","Segment","Segment", "Lympho","Lympho") 
Algorithm = c("CMI", "HiCS","CMI", "HiCS","CMI", "HiCS","CMI", "HiCS", "CMI", "HiCS", "CMI", "HiCS")

AUC <- c(0.953, 0.953, 0.798, 0.800, 0.763, 0.731, 0.956, 0.904, 0.905, 0.825, 0.930, 0.869)
dfAuc = data.frame(Dataset, Algorithm, AUC) 
# barplot(dfAuc$AUC, 
#         names.arg = Dataset,
#         main = "AUC on real world data",
#         xlab="Dataset", 
#         ylab="AUC",
#         ylim = c(0.0, 1.0),
#         col=c("darkblue","deepskyblue"),
#         legend = c("CMI","HiCS"))


library(ggplot2)
ggplot(dfAuc, aes(factor(Dataset), dfAuc$AUC, fill=Algorithm, x=Dataset, y=AUC, gg)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Set2")

``` 

#### Runtime
```{r, echo=FALSE}
Runtime <- c(134.25, 120.01, 29.45, 26.43, 146.22, 36.86, 4547.24, 385.69, 902.45, 122.51, 26.6, 18.76)
dfRun = data.frame(Dataset, Algorithm, Runtime) 
barplot(dfRun$Runtime, 
        names.arg = Dataset,
        main = "Runtime in sec.",
        xlab="Dataset", 
        ylab="Sec.",
        ylim = c(0, 5000),
        col=c("darkblue","deepskyblue"),
        legend = c("CMI","HiCS"))
``` 

#### Subspaces
```{r, echo=FALSE}

SubspaceSizes = c(5, 3.2, 6.2, 2.6, 7.2, 2.4, 6.4, 3.2, 8.2, 2.4, 6.6, 2.2) 
df = data.frame(Dataset, Algorithm, SubspaceSizes) 
# barplot(df$SubspaceSizes, 
#         names.arg = Dataset,
#         main = "Average size of Top5-Subspaces",
#         xlab="Dataset", 
#         ylab="Average Subspace-Size",
#         col=c("darkblue","deepskyblue"),
#         legend = c("CMI","HiCS"))

ggplot(df, aes(factor(Dataset), SubspaceSizes, fill=Algorithm, x=Dataset, y=SubspaceSizes, gg)) +
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Set2") +
  ggtitle("Average size of Top5-Subspaces")

```

With CMI Subspaces with more dimensions were found. 
Some examples of found subspaces:
```{r, echo=FALSE}

library(knitr)

colum <- c(1,2,16:20)
colNam <- c("Algorithm","Dataset","Subspace 1","Subspace 2","Subspace 3","Subspace 4","Subspace 5")

load("results/experimentResult_glass_161013.RData")
kable(combinedResult[which(!duplicated(combinedResult[,16:20])),colum], caption = "Top-Subspaces - Glass-Dataset",
      col.names = colNam)

load("results/experimentResult_ion_161019.RData")
kable(combinedResult[which(!duplicated(combinedResult[,16:20])),colum], caption = "Top-Subspaces - Ion-Dataset",
      col.names = colNam)

load("results/experimentResult_pen_161019.RData")
kable(combinedResult[which(!duplicated(combinedResult[,16:20])),colum], caption = "Top-Subspaces - Pendigits-Dataset",
      col.names = colNam)
```
